COLLISION RATES OF PLANETESIMALS NEAR MEAN-MOTION RESONANCES

### 2020-05-16 main.tex
Outline rest of paper, add discussion of varying mass

We begin by examining simulations $e_{lo}$, $n$ and $e_{hi}$. The positions of the planetesimals in the x-y plane after 5,000 years of integration is shown in figure \ref{fig:xy}. In all cases, the coordinate system is rotated so that the giant planet lies at $\theta = 0$ and the longitude of perihelion of the planet is shown by the dashed line. Resonances with the perturbing planet are visible via axisymmetric gaps. Upon close inspection of the last few simulation snapshots, these gaps appear to follow the planet in its orbit, rather than aligning themselves with the longitude of perihelion. A similar substructure reveals itself in \citet{2000Icar..143...45R} (see bottom panel of figure 3) and \citet{2016ApJ...818..159T} (figure 3). It is worth noting that \citet{2000Icar..143...45R} started with a completely cold planetesimal disk and a Jupiter mass planet on a circular, coplanar orbit. The presence of these features seems robust to the choice of initial conditions.

@@

This also suggests that the features seen in figure \ref{fig:coll_hist_r} are being produced by bodies outside of resonance. To further understand this, we revisit the axisymmetric structure seen in figure \ref{fig:xy}. In figure \ref{fig:coll_polar_e}, we compare the radial collision profile to the polar structure seen in the $e_{low}$, $n$ and $e_{hi}$ simulations. In the lowest forced eccentricity case, the pileups near the edges of the resonances line up with the boundaries of the dip feature seen in the radial collision profile. This is especially prominent for the 2:1 resonance. As discussed previously, the pileups and gaps seen in polar coordinates follow the position of the giant planet in its orbit.

Although the axisymmetric pileup of bodies near the edges of resonances has been seen previously \citep{2000Icar..143...45R, 2016ApJ...818..159T}, we could not find a satisfactory explanation for why it happens and will provide one here. As mentioned previously, the circulation frequency of the critical angle slows as one approaches the edge of the resonance from the outside. Following the pendulum analogy, this is equivalent to reaching the point where the pendulum becomes suspended in the vertical position. When the critical angle stays relatively stationary during an encounter, the net torque will drive the longitude of pericenter towards the point of conjunction \citep{1976ARA&A..14..215P}.  This is the basic mechanism that causes isolated resonance to be stable. Inside of resonance, the critical angles librates about some value and this configuration is maintained. Just outside of resonance, however, the critical angle slowly drifts away, with the drift direction changing sign across the resonance. The net result of this is that bodies near the edge of the resonance experience a torque from that planet that temporarily aligns their pericenters with the orbital phase of the planet. After the encounter ends, differential rotation of the disk slowly destroys the configuration over part of an orbit. This explains why the concentration of planetesimals appears to 'follow' the the planet.

In a disk with no secular forcing, this dynamical phenomena would produce a 'dip' near the center of the resonance and a 'bump' at each edge in cylindrical distance space. When a forced eccentricity is introduced, the edges of the resonance, where the pileups are located, are able to explore a wider range of radii. If the the aphelion distance of the inner edge and the perihelion distance of the outer edge cross, we should expect to see the central dip feature in the collision profile start to disappear. This is exactly what appears to be happening in the middle panel of figure \ref{fig:coll_polar_e}. As the forced eccentricity is increased further, a region forms at the center of the resonance where planetesimals from both sides of the resonance spend time (although not simultaneously). When this occurs, a bump feature forms in the collision profile at this region. This is apparent in the bottom panel of figure \ref{fig:coll_polar_e}.

@@

\subsection{Collision Rates}\label{sec:coll_rates}

\subsection{Where Does the Dust End Up?}

\section{Influence of Jupiter's Mass and Eccentricity on the Radial Dust Profile}

\section{Observability of Dust} \label{sec:dust}

Aaron's CASA images go here

\section{Summary and Discussion} \label{sec:discuss}

>>>>

We begin by examining simulations $e_{lo}$, $n$ and $e_{hi}$. The positions of the planetesimals in the x-y plane after 5,000 years of integration is shown in figure \ref{fig:xy}. In all cases, the coordinate system is rotated so that the giant planet lies at $\theta = 0$ and the longitude of perihelion of the planet is shown by the dashed line. Resonances with the perturbing planet are visible via nonaxisymmetric gaps. Upon close inspection of the last few simulation snapshots, these gaps appear to follow the planet in its orbit, rather than aligning themselves with the longitude of perihelion. A similar substructure reveals itself in \citet{2000Icar..143...45R} (see bottom panel of figure 3) and \citet{2016ApJ...818..159T} (figure 3). It is worth noting that \citet{2000Icar..143...45R} started with a completely cold planetesimal disk and a Jupiter mass planet on a circular, coplanar orbit. The presence of these features seems robust to the choice of initial conditions.

@@

This also suggests that the features seen in figure \ref{fig:coll_hist_r} are being produced by bodies outside of resonance. To further understand this, we revisit the nonaxisymmetric structure seen in figure \ref{fig:xy}. In figure \ref{fig:coll_polar_e}, we compare the radial collision profile to the polar structure seen in the $e_{low}$, $n$ and $e_{hi}$ simulations. In the lowest forced eccentricity case, the pileups near the edges of the resonances line up with the boundaries of the dip feature seen in the radial collision profile. This is especially prominent for the 2:1 resonance. As discussed previously, the pileups and gaps seen in polar coordinates follow the position of the giant planet in its orbit.

Although the nonaxisymmetric pileup of bodies near the edges of resonances has been seen previously \citep{2000Icar..143...45R, 2016ApJ...818..159T}, we could not find a satisfactory explanation for why it happens and will provide one here. As mentioned previously, the circulation frequency of the critical angle slows as one approaches the edge of the resonance from the outside. Following the pendulum analogy, this is equivalent to reaching the point where the pendulum becomes suspended in the vertical position. When the critical angle stays relatively stationary during an encounter, the net torque will drive the longitude of pericenter towards the point of conjunction \citep{1976ARA&A..14..215P}.  This is the basic mechanism that causes isolated resonance to be stable. Inside of resonance, the critical angles librates about some value and this configuration is maintained. Just outside of resonance, however, the critical angle slowly drifts away, with the drift direction changing sign across the resonance. The net result of this is that bodies near the edge of the resonance experience a torque from that planet that temporarily aligns their pericenters with the orbital phase of the planet. After the encounter ends, differential rotation of the disk slowly destroys the configuration over part of an orbit. This explains why the concentration of planetesimals appears to 'follow' the the planet.

In a disk with no secular forcing, this dynamical phenomena would produce a 'dip' near the center of the resonance and a 'bump' at each edge in cylindrical distance space. When a forced eccentricity is introduced, the edges of the resonance (indicated by the solid and dashed lines in figure \ref{fig:coll_polar_e}), where the pileups are located, begin to explore a wider range of radii. If the the aphelion distance of the inner edge and the perihelion distance of the outer edge cross, we should expect to see the central dip feature in the collision profile disappear. This is exactly what appears to be happening in the middle panel of figure \ref{fig:coll_polar_e}. As the forced eccentricity is increased further, a region forms at the center of the resonance where planetesimals from both sides of the resonance spend time (although not simultaneously). When this occurs, a bump feature forms in the collision profile near the center of the resonance. This is apparent in the bottom panel of figure \ref{fig:coll_polar_e}.

Although the same phenomenon appears to be happening near the 3:1 MMR, the pileups at the edges of the resonance are not as well defined. We attribute this to the longer synodic period, compared to the 2:1 MMR. In addition, a number of other nearby resonances, including the 7:2 (at 2.25 AU), the 10:3 (at 2.33 AU), the 8:3 (at 2.71 AU) and the 5:3 (at 2.82 AU) also likely are contributing to the collision profile in cylindrical distance space near the 3:1. The width of the 3:1 MMR is also much narrower than the 2:1, which means that any dip or bump features produced by it would require much higher resolution observations. For these reasons, we focus on the 2:1 MMR as the main diagnostic indicator.

\subsection{Varying the Mass}

The dynamical effects on the planetesimal disk of varying the mass of the planet of varying the mass are much simpler, in that it doing so does not affect the forced eccentricity. The width of a first order resonance scales with $m$ (see equation \ref{eq:res_so}), while the leading order terms in the resonant part of the disturbing function, which set the strength of the resonance, also scale as $m$. For the 2:1 MMR, this dynamics near resonance are equally sensitive to changes in eccentricity and mass.

We show the polar structure of the $m_{low}$, $n$ and $m_{hi}$ simulations in figure \ref{fig:coll_polar_m} alongside the radial collision profile. In all three cases, the eccentricity of the perturbing planet is set to $e_{jup}$ and so the forced eccentricity does not change. Changes in the apocenter and pericenter distances of the edges of the resonances are entirely due to changes in the libration width of the resonances here. For the $m_{low}$ and $n$ simulations, the inner apocenter and outer pericenter distances near the 2:1 MMR are quite similar and no strong features appear in the collision profile near this region. For the $m_{hi}$ case, the edges of the resonances are sufficiently far apart to allow a gap near the center of the 2:1 resonance to appear in the collision profile. As seen in the previous section, the collision profile does not appear to follow the simple dip vs bump behavior.

\section{Constraining the Mass and Eccentricity of the Planet}

\subsection{Simulated ALMA Observations}

Aaron's stuff goes here. Can the bumps and dips near the 2:1 MMR actually be resolved? What does the 3:1 resonance look like? Although we might not be able to use the 3:1 MMR to constrain mass or eccentricity, at least two resonances must be identified to determine where the 2:1 MMR actually is.

\subsection{Implications of a Bump vs Dip at the 2:1 MMR}

Show figure \ref{fig:bump_dip_diag} here. The presence of a dip or a bump at the 2:1 MMR can be used to determine which side of the eccentricity-mass plane we are on. Resonance strength scales with $m$ and $e$, so this will generally work better up and to the right on this diagram. Parameter combinations that lie between the dashed lines mean that the inner apo - outer peri distance is smaller than the resonance width. In this case, there won't really be much of a feature.

\section{Discussion and Summary}

@@

\begin{figure*}
\begin{center}
    \includegraphics[width=\textwidth]{figures/bump_dip_diag.png}
    \caption{The presence of a dip or bump near the resonances can be used to constrain the mass and eccentricity
    of the perturbing planet. The color scale indicates the relative strength of the features, while the dashed lines
    indicate boundaries in parameter space where a dip or a bump will be produced at the resonance.\label{fig:bump_dip_diag}}
\end{center}
\end{figure*}
